Topological Equivariant Artist Model for Visualization Library Architecture

### 2024-05-20 paper/main.tex
sketch of interactivity

The structure on values has traditionally been codified using the measurement scales-i.e. nominal, ordinal, interval, ratio - which Steven's classified by their mathematical structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, homasMathematizationNotMeasurement2014}.

@@

A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
\begin{enumerate}
  \item all fibers in the bundle are isomorphic. Since all fibers are isomorphic, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
  %the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$. All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$.
  \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial. \textbf{Local triviality} means $\dtotal\restriction_{\openset} = \openset\times F$. In this paper we use $\dtotal\restriction_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset, and a \textbf{local trivialization} is a specific choice of neighborhoods and their preimages.
  %there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022} defined in \texttt{amssymb}. For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ defined only on the points in $\openset$. Since $\pi^{-1}(\dbase)=\dtotal$, we use the shorthand $\dtotal\restriction_{\openset} \coloneqq \pi^{-1}\restriction_{\openset}$}, is locally trivial. The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
\end{enumerate}
\end{definition}
\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}

@@

@@

While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.
\begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
\begin{itemize}
  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|_{\openset} =  \dsection^{b}|_{\openset}$.
  \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
  \end{itemize}
\end{definition}
The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections. The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values. The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections. Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (indexing subspaces) is preserved.

@@

\paragraph{\textbf{Graphic Corresponding to Data}}
The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.
\begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as
  \begin{equation*}
    \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
  \end{equation*}
for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
\end{definition}
The pushforward sheaf returns the set of graphic sections over the data base space that corresponds to the graphic space $\vindex^{-1}(\openset) = \opensetgc$. The pushforward functor $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc}
\end{equation}
such that it provides a way to look up which graphic corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
\end{equation}
such that $\vindexpush\gsection(\dbasepoint))(\gbasepoint) = \gsection(\gbasepoint)$ for all $\gbasepoint \in \vindexpre(\dbasepoint)$. Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.

\paragraph{\textbf{Data Corresponding to Graphic}}
The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space.
\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
  \begin{equation*}
    \vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))
  \end{equation*}
for $\vindex(\opensetgc) \in \dbase$.
\end{definition}
The pullback sheaf returns the set of data sections over the graphic base space that corresponds to the graphic space $\vindex(\opensetg) = \openset$. The pullback $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
\begin{equation}
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc}
\end{equation}
such that there is a way to then look up what data values correspond with a graphic index
\begin{equation}
  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
\end{equation}
As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.

@@

>>>>

The structure on values has traditionally been codified using the measurement scales-i.e. nominal, ordinal, interval, ratio - which Steven's classified by their mathematical structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.

@@

A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
\begin{enumerate}
  \item all fibers in the bundle are isomorphic. Since all fibers are isomorphic, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
  %the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$. All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$.
  \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial. \textbf{Local triviality} means $\dtotal\restriction_{\openset} = \openset\times F$. In this paper we use $\dtotal\restriction_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset, and a \textbf{local trivialization} is a specific choice of neighborhoods and their preimages.
  %there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022} defined in \texttt{amssymb}. For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ defined only on the points in $\openset$. Since $\pi^{-1}(\dbase)=\dtotal$, we use the shorthand $\dtotal\restriction_{\openset} \coloneqq \pi^{-1}\restriction_{\openset}$}, is locally trivial. The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
\end{enumerate}
\end{definition}

@@

\note{ending on sectionsis probably better for our purposes!}
\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}

@@

While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.
\begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
\begin{itemize}
  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|_{\openset} =  \dsection^{b}|_{\openset}$.
  \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
  \end{itemize}
\end{definition}
The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections. The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values. The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections. Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (indexing subspaces) is preserved.

@@

\paragraph{\textbf{Graphic Corresponding to Data}}
The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.
\begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as
  \begin{equation}
    \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
  \end{equation}
for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
\end{definition}
The pushforward sheaf returns the set of graphic sections over the data base space that corresponds to the graphic space $\vindex^{-1}(\openset) = \opensetgc$. The pushforward functor $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}
  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc}
\end{equation}
such that it provides a way to look up which graphic corresponds with a data index
\begin{equation}
  \label{eq:atct:sheaf:pushforward_select}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
\end{equation}
such that $\vindexpush\gsection(\dbasepoint))(\gbasepoint) = \gsection(\gbasepoint)$ for all $\gbasepoint \in \vindexpre(\dbasepoint)$. Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.

\paragraph{\textbf{Data Corresponding to Graphic}}
The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space.
\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
  \begin{equation*}
    \vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))
  \end{equation*}
for $\vindex(\opensetgc) \in \dbase$.
\end{definition}
The pullback sheaf returns the set of data sections over the graphic base space that corresponds to the graphic space $\vindex(\opensetg) = \openset$. The pullback $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
\begin{equation}
  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc}
\end{equation}
such that there is a way to then look up what data values correspond with a graphic index
\begin{equation}
  \label{eq:atct:sheaf:pullback_hover}
  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
\end{equation}
As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.

@@

\note{explain why annotation is similar to brush/linking in operators section }

\subsection{Animation and Interactivity}

\begin{LaTeXdescription}
  \item[pan, zoom, scroll] sheaf: locality + gluing \autoref{def:atct:sheaf}
  \item[selection and hover] pushforward \autoref{eq:atct:sheaf:pushforward_select},  pullback \autoref{eq:atct:sheaf:pullback_hover}
  \item[brushing, linking, annotation] composition of artists \autoref{eq:artist:join_base_fiber}
\end{LaTeXdescription}

Animation and interaction are a set of stills. Because the constraints are on the functions $\vartist\circ\dsection$ , satisfying the constraints on each function means that the constraint is satisfied for all visualizations $\{\vartist(\dsection(\dbasepoint)) \mid\dbasepoint \in \dbase \}$ that make up an animation or interaction.
